Reproducible development environments with rix

### 2023-12-23 paper.qmd
github actions

Obviously, writing such an expression will be quite complicated for someone that
is unfamiliar with the [Nix]{.proglang} programming language, but the prospect
of being able to define a completely reproducible development environment in a
single text file and then use that file to build and use that environment
anywhere is quite interesting. Because we thought this, we wrote the [rix]{.pkg}
which comes with the `rix()` function that helps generate such expressions quite
easily. For example, here is how one could use [rix]{.pkg} to generate a
`default.nix` file that installs version 4.3.1 of [R]{.proglang} and the
[dplyr]{.pkg} and [chronicler]{.pkg} packages (we don't show the generated
expression here because of space):

@@

@@

```{r}
rix(r_ver = "4.3.1",
    r_pkgs = c("dplyr", "chronicler", "AER@1.2-8"),
    system_pkgs = c("quarto", "git"),
    tex_pkgs = c(
          "amsmath",
          "framed",
          "fvextra",
          "environ",
          "fontawesome5",
          "orcidlink",
          "pdfcol",
          "tcolorbox",
          "tikzfill"
    ),
    git_pkgs = list(
                 list(package_name = "rix",
                   repo_url = "https://github.com/b-rodrigues/rix/",
                   branch_name = "master",
                   commit = "ea92a88ecdfc2d74bdf1dde3e441d008521b1756"),
                 list(package_name = "fusen",
                   repo_url = "https://github.com/ThinkR-open/fusen",
                   branch_name = "main",
                   commit = "d617172447d2947efb20ad6a4463742b8a5d79dc")
    ),
    ide = "rstudio",
    project_path = tempfile(),
    overwrite = TRUE)
```

>>>>

Obviously, writing such an expression will be quite complicated for someone that
is unfamiliar with the [Nix]{.proglang} programming language, but the prospect
of being able to define a completely reproducible development environment in a
single text file and then use that file to build and use that environment
anywhere is quite interesting. Because we believed that lowering the cost of
writing Nix expressions would lead to a wider adoption of the Nix package
manager for reproducibility, we wrote the [rix]{.pkg} which comes with the
`rix()` function that helps generate such expressions quite easily. For example,
here is how one could use [rix]{.pkg} to generate a `default.nix` file that
installs version 4.3.1 of [R]{.proglang} and the [dplyr]{.pkg} and
[chronicler]{.pkg} packages (we don't show the generated expression here because
of space):

@@

[rix]{.pkg} also provides installation intructions for Nix, but [rix]{.pkg} can
generate Nix expression even if Nix is not installed on one's system. This is
quite useful if one wants to generate an expression and use that expression in a
CI/CD environment such as Github Actions. For example, the repository that
contains the source code this article^[https://github.com/b-rodrigues/rix_paper]
uses Github Actions for compilation. Essentially, each time one of the authors
pushed to the master branch, a Github Actions runner would be started, Nix would
get installed on said runner, and an environment containing the right packages
was generated from a `default.nix` file that is hosted on the same repository.
The paper would then get compiled in this reproducible environment.

@@

```{r}
library(rix)

rix(r_ver = "4.3.1",
    r_pkgs = c("dplyr", "chronicler", "AER@1.2-8"),
    system_pkgs = c("quarto", "git"),
    tex_pkgs = c(
          "amsmath",
          "framed",
          "fvextra",
          "environ",
          "fontawesome5",
          "orcidlink",
          "pdfcol",
          "tcolorbox",
          "tikzfill"
    ),
    git_pkgs = list(
                 list(package_name = "rix",
                   repo_url = "https://github.com/b-rodrigues/rix/",
                   branch_name = "master",
                   commit = "ea92a88ecdfc2d74bdf1dde3e441d008521b1756"),
                 list(package_name = "fusen",
                   repo_url = "https://github.com/ThinkR-open/fusen",
                   branch_name = "main",
                   commit = "d617172447d2947efb20ad6a4463742b8a5d79dc")
    ),
    ide = "rstudio",
    project_path = tempfile(),
    overwrite = TRUE)
```
